---
title: "SMcLaughlin ST537 R Project"
author: "Sarah McLaughlin"
date: "10/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(klaR)
library(MASS)
library(knitr)
```

# Classification Method for Analysis of Wine Data Set  
Bring in the data and split into training and testing sets.  
```{r SM wine data}
# Bring in the two data sets 
red.wine <- read.csv("winequality-red.csv", header = TRUE, sep = ";")
white.wine <- read.csv("winequality-white.csv", header = TRUE, sep = ";")

# Create red and white variable  
red.wine$color <- c("red")
white.wine$color <- c("white")

# Combine data into one data set  
wine <- rbind(red.wine, white.wine)
# Make color a factor 
wine$color <- as.factor(wine$color)

# Train and test set 
set.seed(1)
train <- sample(1:nrow(wine), size = nrow(wine)*0.8)
test <- setdiff(1:nrow(wine), train)

wine.train <- wine[train, ]
wine.test <- wine[test, ]
```  

Calculate the "No Information Rate". 
```{r SM NIR}
# Calculate frequency table
tab <- table(wine$quality)
tab
# Calculate NIR
NIR <- max(tab)/sum(tab)
```
The No Information Rate is `r round(NIR, 3)`. Thus, if we classified each wine sample into the largest quality group (6), our model would be correct about 44% of the time. The Classification Model that we create must have a better accuracy than 44%.  

## Linear Discriminant Analysis  
**Assumptions:**  

**Running the Linear Discriminant Analysis:**  
```{r SM LDA}
# Create LDA Model with Training Set 
lda.out <- lda(quality ~ ., data = wine.train)

# Calculate test quality using LDA Model 
lda.pred <- predict(lda.out, newdata = wine.test)

# Create error matrix  
lda.err <- errormatrix(true = wine.test$quality, predicted = lda.pred$class, 
                       relative = TRUE)

kable(round(lda.err, 3))
```

